---
title: "sim-validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sim-validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```


```{r setup}
library("splithalf")
library("tidyr")

```

# Usage

This script runs _splithalf_ with many possible combinations of the options. Its main use is to ensure that the _splithalf_ function 

We want to run the following:

outcome x score x halftype x average = 3x3x2 = 18

* outcome = "RT"
* outcome = "accuracy"

* score = "average"
* score = "difference" (the default option)
* score = "difference_of_difference"

* halftype = "oddeven"
* halftype = "halfs"
* halftype = "random"

* average = "mean"
* average = "median"


```{r generate_data}
# Thanks to Nathaniel Haines for the suggestion and the starting code (https://github.com/sdparsons/splithalf/issues/9). 

set.seed(01042021)

n_participants <- 60 ## sample size
n_trials <- 80
n_blocks <- 2

sim_data <- data.frame(participant_number = rep(1:n_participants, each = n_blocks * n_trials),
                       trial_number = rep(1:n_trials, times = n_blocks * n_participants),
                       block_name = rep(c("A","B"), 
                                        each = n_trials, 
                                        length.out = n_participants * n_trials * n_blocks),
                       trial_type = rep(c("congruent","congruent",
                                          "incongruent","incongruent"),
                                        length.out = n_participants * n_trials * n_blocks / 2),
                       RT = rnorm(n_participants * n_trials * n_blocks, 500, 200),
                       ACC = 1)

sim_data_acc <- data.frame(participant_number = rep(1:n_participants, each = n_blocks * n_trials),
                       trial_number = rep(1:n_trials, times = n_blocks * n_participants),
                       block_name = rep(c("A","B"), 
                                        each = n_trials, 
                                        length.out = n_participants * n_trials * n_blocks),
                       trial_type = rep(c("congruent","congruent",
                                          "incongruent","incongruent"),
                                        length.out = n_participants * n_trials * n_blocks / 2),
                       RT = rnorm(n_participants * n_trials * n_blocks, 500, 200),
                       ACC = c(rbinom(n_participants *
                                    n_trials *
                                    n_blocks / 6, 
                                    1, .5),
                               rbinom(n_participants *
                                    n_trials *
                                    n_blocks / 6, 
                                    1, .7),
                               rbinom(n_participants *
                                    n_trials *
                                    n_blocks / 6, 
                                    1, .9),           
                               rbinom(n_participants *
                                    n_trials *
                                    n_blocks / 6, 
                                    1, .5),
                               rbinom(n_participants *
                                    n_trials *
                                    n_blocks / 6, 
                                    1, .7),
                               rbinom(n_participants *
                                    n_trials *
                                    n_blocks / 6, 
                                    1, .9)))




```


```{r test_setup}
params <- list(score = c("average",
                         "difference",
                         "difference_of_difference"),
               halftype = c("oddeven",
                            "halfs",
                            "random"),
               averae = c("mean",
                          "median") )

expand.grid(params) %>%
  dplyr::arrange(., score, halftype)

```

# Response times

```{r rt_test1}
splithalf(data = sim_data,
          outcome = "RT",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r rt_test2}
splithalf(data = sim_data,
          outcome = "RT",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r rt_test3}
splithalf(data = sim_data,
          outcome = "RT",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r rt_test4}
splithalf(data = sim_data,
          outcome = "RT",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r rt_test5}
splithalf(data = sim_data,
          outcome = "RT",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean",
          plot = TRUE)
```

```{r rt_test6}
splithalf(data = sim_data,
          outcome = "RT",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median",
          plot = TRUE)
```

```{r rt_test7}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r rt_test8}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r rt_test9}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r rt_test10}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r rt_test11}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean",
          plot = TRUE)
```

```{r rt_test12}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median",
          plot = TRUE)
```

```{r rt_test13}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r rt_test14}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r rt_test15}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r rt_test16}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r rt_test17}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean",
          plot = TRUE)
```

```{r rt_test18}
splithalf(data = sim_data,
          outcome = "RT",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median",
          plot = TRUE)
```

# Average accuracy rates

```{r acc_test1}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r acc_test2}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r acc_test3}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r acc_test4}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r acc_test5}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean",
          plot = TRUE)
```

```{r acc_test6}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "average",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median",
          plot = TRUE)
```

```{r acc_test7}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r acc_test8}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r acc_test9}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r acc_test10}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r acc_test11}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean",
          plot = TRUE)
```

```{r acc_test12}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median",
          plot = TRUE)
```

```{r acc_test13}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r acc_test14}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "oddeven",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r acc_test15}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean")
```

```{r acc_test16}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "halfs",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median")
```

```{r acc_test17}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "mean",
          plot = TRUE)
```

```{r acc_test18}
splithalf(data = sim_data_acc,
          outcome = "accuracy",
          score = "difference_of_difference",
          conditionlist = c("A", "B"),
          halftype = "random",
          permutations = 500,
          var.RT = "RT",
          var.condition = "block_name",
          var.participant = "participant_number",
          var.compare = "trial_type",
          var.ACC = "ACC",
          compare1 = "congruent",
          compare2 = "incongruent",
          average = "median",
          plot = TRUE)
```

